\chapter*{Acknowledgements}

This thesis marks the culmination of my journey in the MSc Geomatics program, which would not have been possible without the support and guidance of many individuals and institutions.

First and foremost, I would like to express my sincere gratitude to my primary supervisor, Dr. Hugo Ledoux, for his exceptional guidance, unwavering patience, and continuous encouragement throughout this research. His deep expertise in 3D city models and geospatial data formats has been invaluable, and his critical insights significantly shaped the direction and quality of this work.

I am equally grateful to my secondary supervisor, Dr. Ravi Peters, who made preliminary work on this project and whose pragmatic approach to problem-solving helped me tackle numerous challenges in the development of FlatCityBuf. His constructive feedback consistently pushed me to refine both my ideas and their implementations.

My heartfelt thanks extend to Dr. Martijn Meijers, who served as co-reader for this thesis. His thoughtful comments and perspectives contributed substantially to improving the final manuscript.

I would like to acknowledge Delft University of Technology, particularly the Faculty of Architecture and the Built Environment and the MSc Geomatics program, for providing an exceptional academic environment and resources. The administrative and technical staff deserve special mention for their consistent support throughout my studies.

My journey would not have been complete without my colleagues in the Geomatics program. Their camaraderie, intellectual discussions, and mutual support created a stimulating environment that fostered both academic growth and lasting friendships.

I am also indebted to the FlatGeoBuf team, whose pioneering work in cloud-optimized geospatial formats laid important groundwork for this research. Their open-source contributions and technical insights have been inspirational.

Beyond academia, I am grateful to my colleagues at Eukarya Inc., who showed remarkable understanding and flexibility as I balanced professional responsibilities with academic pursuits. Their support and accommodation made it possible for me to pursue both paths simultaneously.

My deepest appreciation goes to my family for their unconditional love, encouragement, and belief in me. Their unwavering support has been my anchor throughout this journey, especially during challenging periods.

Finally, a special acknowledgment to Chiharu, my partner, whose patience, understanding, and constant encouragement have been my greatest source of strength.
\begin{CJK}{UTF8}{min}
  智春、研究が大変なときも疲れたときもいつも陰ながら応援してくれて本当にありがとう。智春の支えがあって、今日まで頑張ってくることができました。
\end{CJK}

To everyone who has been part of this journey—thank you.

\vspace{1cm}

\begin{flushright}
  \textit{Hidemichi Baba}\\
  Delft, \today
\end{flushright}

\ldots
